

---
title: "R Notebook"
output:
  html_notebook: default
  pdf_document: default
---
## Project 1: ACS Quality of Life per State, Education and Gender

We want to find a relationship betweeen the Attained Education, State, Gender, and the quality of life of US population.

The main assumption is that quality of life is directly proportional to with Salary [USD/hour] and Commute Time to work [hours].


## Load the libraries and data

```{r, message=F}
library(dplyr)
library(readr)
library(DT)
library(plotly)
library(ggplot2)
library(RColorBrewer)
library(d3heatmap)

load.data <- FALSE
```

```{r, include=F}
if(load.data){
  acs14pusa <- read_csv("./data/ss14pusa.csv", guess_max = 10000)
  acs14pusb <- read_csv("./data/ss14pusb.csv", guess_max = 10000)

} else {
  acs14pus <- read_csv(file = "./output/ss14pus_columns.csv" )
}


```

```{r}
if(load.data){
  datatable(head(acs14pusa,50), options = list(scrollX=T, pageLength = 10))
}
```

## Basic information about original data
```{r,message=F}
if(load.data){
  dim(acs14pusa)
  dim(acs14pusb)
}
```

## Select relevant columns to reduce memory usage.
```{r,message=F}
if(load.data){
  relevant.columns <- c("SERIALNO", "ST", "SEX", "AGEP", "SCHL", "INDP", "WKHP", "WAGP", "ESR",
                        "PINCP", "PERNP", "JWMNP")
  acs14pusa.cols <- acs14pusa[,colnames(acs14pusa)%in%relevant.columns]
  acs14pusb.cols <- acs14pusb[,colnames(acs14pusb)%in%relevant.columns]
  
  rm(acs14pusa, acs14pusb)
  gc()
}
```

## Bind part a and b, and get basic information
```{r,message=F}
if(load.data){
  acs14pus <- rbind(acs14pusa.cols, acs14pusb.cols)
  dim(acs14pus)
}

```

## Add state names and abbreviations
Credits to Arnold Chua Lau (Spring 2016).

```{r, message=F}
if(load.data){  
  ST.anno = read_csv("./data/statenames.csv")
  ST.anno = mutate(ST.anno, STabbr=abbr, STname=name)
  
  acs14pus = mutate(acs14pus, STnum = as.numeric(ST))
  acs14pus <- left_join(acs14pus, ST.anno, by = c("STnum" = "code"))
  
  select(sample_n(acs14pus,5), starts_with("ST"))
}
```

## Convert data

```{r, message=F}

acs14pus$JWMNP <- as.numeric(acs14pus$JWMNP)
acs14pus$WAGP <- as.numeric(acs14pus$WAGP)
acs14pus$WKHP <- as.numeric(acs14pus$WKHP)
acs14pus$STabbr <- as.factor(acs14pus$STabbr)

acs14pus$WAGEHOUR <- acs14pus$WAGP / acs14pus$WKHP / 52

```


## Write result to a csv file. So we do not need to build them again (takes 30 min in a laptop)
```{r,message=F}
if(load.data){
  write_csv(x = acs14pus, path = "./output/ss14pus_columns.csv" )
}
```


# Summary statistics (Jaime)

# Summary Statistics
```{r, message=FALSE}
summary.mean <- acs14pus %>% group_by(STabbr) %>% summarise(mean(na.omit(JWMNP)),
                                                       mean(na.omit(WAGP)),
                                                       mean(na.omit(WKHP)),
                                                       mean(na.omit(WAGEHOUR))
                                                       )

summary.mean[,-1] <- round(summary.mean[,-1],1)
names(summary.mean) <- c("STabbr", "JWMNP", "WAGP", "WKHP", "WAGEHOUR")

datatable(summary.mean, options = list(scrollX=T, pageLength = length(summary.mean$STabbr)))

```


# BoxPlot - Commute Time
```{r, message=FALSE, echo=FALSE}

gc()

summary.mean <- summary.mean[order(summary.mean$JWMNP),]
acs14pus$STabbr <- factor(acs14pus$STabbr, levels = summary.mean$STabbr)

plot_ly(x = acs14pus$STabbr , y = acs14pus$JWMNP , type = "box") %>%
  layout(title = "Commute time per State",
         xaxis = list(title ="State"),
         yaxis = list(title = "Commute time (minutes)")
         )



```

# BoxPlot - Wage per Hour (Jaime)
# JG: I am not convinced of these data values. DC is far too high!

```{r, message=FALSE, echo=FALSE}
plot_ly(x = acs14pus$STabbr , y = acs14pus$WAGP , type = "box") %>%
  layout(title = "12-Month Wage per State",
         scene = list(
           xaxis = list(title ="State"),
           yaxis = list(title = "Wage")
         ))

```


## Heatmap

```{r, message=FALSE}
summary.median.2 <- acs14pus %>% group_by(SCHL, INDP) %>% summarise(median(na.omit(JWMNP)),
                                                       median(na.omit(WAGP)),
                                                       median(na.omit(WKHP)),
                                                       median(na.omit(WAGEHOUR))
                                                       )

summary.median.2[,-c(1,2)] <- round(summary.median.2[,-c(1,2)],1)
names(summary.median.2) <- c("SCHL", "INDP", "JWMNP", "WAGP", "WKHP", "WAGEHOUR")
summary.median.2$SCHL <- as.factor(summary.median.2$SCHL)
summary.median.2$INDP <- as.integer(summary.median.2$INDP)

summary.median.2 <- na.omit(summary.median.2)

industry.categories = read_csv("./data/industry_codes.csv")

summary.median.2 = mutate(summary.median.2, INDP = as.numeric(INDP))
summary.median.2 <- left_join(summary.median.2, industry.categories, by = c("INDP" = "code"))


plot_ly(z = summary.median.2$JWMNP, 
        x = summary.median.2$SCHL,
        y = summary.median.2$industry,
        type = "heatmap") %>%
  layout(title = "Commute time by Education Attainment and Industry",
         xaxis = list(title ="Education"),
         yaxis = list(title = "Industry"),
         width = 1000,
         height = 700
         )


```




## Choropleth maps of US states (Ying)


