## Preprocess data


  Variables     | Dictionary
  ------------- | -------------
1.SERIALNO      | Unique code for person
2.ST            | State Code
3.SEX           | Gender
4.AGEP          | Age
5.SCHL          | Educational Attainment
6.FOD1P         | Field of Degree - First Entry
7.FOD1P         | Field of Degree - Second Entry
8.WKHP          | Hours Worked per Week
9.WAGP          | Wages or salary income
10.ESR          | Employment Status Recode
11.PINCP        | Total Person's Income
12.PERNP        | Total Person's Earning
13.JWMNP        |  Travel time to work


```{r, echo=FALSE}
options(warn=-1)
setwd("~/Desktop/G5243/project1/personal data/2014 all person")
Origdata1<-read.csv("ss14pusa.csv",header=T)
Proj_1a<-cbind(Origdata1$SERIALNO,Origdata1$ST,Origdata1$SEX,
               Origdata1$AGEP,Origdata1$SCHL,Origdata1$FOD1P,Origdata1$FOD2P,
               Origdata1$WKHP,Origdata1$WAGP,Origdata1$ESR,Origdata1$PINCP,
               Origdata1$PERNP,Origdata1$JWMNP)
Origdata2<-read.csv("ss14pusb.csv",header=T)
Proj_1b<-cbind(Origdata2$SERIALNO,Origdata2$ST,Origdata2$SEX,
               Origdata2$AGEP,Origdata2$SCHL,Origdata2$FOD1P,Origdata2$FOD2P,
               Origdata2$WKHP,Origdata2$WAGP,Origdata2$ESR,Origdata2$PINCP,
               Origdata2$PERNP,Origdata2$JWMNP)
Proj_1<-rbind(Proj_1a,Proj_1b)

colnames(Proj_1) <- c("SERIALNO","ST","SEX","AGEP","SCHL",
                      "FOD1P","FOD2P","WKHP","WAGP","ESR",
                      "PINCP","PERNP","JWMNP")
data <- Proj_1
data1 <- Origdata1
data2 <- Origdata2
```

```{r, echo=FALSE}
SERIALNO<-Proj_1[,1]
ST<-Proj_1[,2] #State Code
SEX<-Proj_1[,3]#Gender
AGEP<-Proj_1[,4]# Age
SCHL<-Proj_1[,5]#Educational Attainment
FOD1P<-Proj_1[,6]#Field of Degree - First Entry
FOD2P<-Proj_1[,7]#Field of Degree - Second Entry
WKHP<-Proj_1[,8]#Hours Worked per Week
WAGP<-Proj_1[,9]#Wages or salary income
ESR<-Proj_1[,10]#Employment Status Recode
PINCP<-Proj_1[,11]#Total Person's Income
PERNP<-Proj_1[,12]#Total Person's Earning
JWMNP<-Proj_1[,13]#Travel time to work
```
Here we have to read the dictionary carefully since there are NA in the columns are actually have means.
The meanings in the NA in different variables:

Variables     | NA meanings
------------- | -------------
SERIALNO      | No NA
ST            | No NA
SEX           | No NA
AGEP          | No NA
SCHL          | Less than 3 years old
FOD1P         | Less than bachelor's degree
FOD1P         | Less than bachelor's degree
WKHP          | Less than 16 years old/did not work
WAGP          | Less than 15 years old
ESR           | Less than 16 years old
PINCP         | Less than 15 years old
PERNP         | Less than 15 years old
JWMNP         | not a worker or worker who worked at home


## 1.Income status in US
### 1.1 Gender, Age and Income
```{r,echo=FALSE}
library(ggplot2)
library(plotly)
data1.2 <- data[,c(3,4,11)] 
male <- data1.2[which(data1.2[,1]==1),2:3] 
female <- data1.2[which(data1.2[,1]==2),2:3]
male <- male[which(male[,2]>0),]
female <- female[which(female[,2]>0),]
male_av <- matrix(c(seq(11,91,length=17),rep(0,17)),ncol=2)
female_av <- matrix(c(seq(11,91,length=17),rep(0,17)),ncol=2)
for (i in 1:17)
{
  male_av[i,2] <- mean(male[which(male[,1]>5*i+5 & male[,1]<5*i+11),2])
  female_av[i,2] <- mean(female[which(female[,1]>5*i+5 & female[,1]<5*i+11),2])
}
male_av1 <- cbind(rep(1,17),male_av)
female_av1 <- cbind(rep(2,17),female_av)
final <- rbind(male_av1,female_av1)
final <- as.data.frame(final)
final[1:17,1] <- "male"
final[18:34,1] <- "female"
names(final) <- c("Gender", "Age", "Income")
```
```{r}
plot_ly(data = final, x = Age, y = Income, mode = "markers", 
        color = final[,1], size = final[,2], title="a")%>%
  layout(title = '2014 US averge personal income by Gender and Age')
```


### 1.2 Education, Working Hours and Income
```{r,echo=FALSE}
data1.2<- data[,c(5,8,11)]
data1.2 <- data1.2[which(rowSums(data1.2)>0),]
INCOME <- matrix(0,nrow=8,ncol=10)
for(i in 1:8){
  for(j in 1:10){
    index <- which(data1.2[,1]>3*i-3 & data1.2[,1]<3*i+1 & data1.2[,2]>10*j-10 & data1.2[,2]<10*j+1)
    INCOME[i,j] <- mean(data1.2[index,3])
  }
}
```
```{r}
plot_ly(z=INCOME,type="surface",colorbar = list(title = "USD"))%>%
  layout(title = '2014 US averge personal income by Education and Working Hours', 
         xaxis = list(title='EDU'))
```


### 1.3 Occupation and income
```{r,echo=FALSE}
subdata1.5 <- cbind(c(data1$INDP,data2$INDP),data[,11])
subdata1.5 <- subdata1.5[which(rowSums(subdata1.5)>0),]
ind_code <- read.csv("~/Desktop/industry_codes.csv")
index <- rep(0,267)
for(i in 1:267)
{
  index[i] <- which(unique(ind_code[,2])==ind_code[i,2])
}
ind_code <- cbind(index,ind_code)
subdata1.5 <- cbind(rep(0,nrow(subdata1.5)),subdata1.5)
for(i in 1:19)
{
  subdata1.5[which(subdata1.5[,2] %in% ind_code[which(ind_code[,1]==i),2]),1] <- i
}
INDincome <- rep(0,19)
for(i in 1:19)
{
  INDincome[i] <- mean(subdata1.5[which(subdata1.5[,1]==i),3])
}
```
```{r}
barplot(sort(INDincome), horiz=T,col = "light blue",xlab = "average income",
        main = "Industry Income", border=F,xlim=c(0,80000),
        cex.axis=0.7,names.arg = unique(ind_code[,3])[order(INDincome)],
        cex.names=0.6,las=1)
```


### 1.4 States and income
```{r,echo=FALSE}
subdata1 <- data[,c(2,11)]
subdata1 <- subdata1[which(subdata1[,2]>0),]
STcode <- c(1,2,4,5,6,8,9,10,12,13,15:42,44:51,53:56)

STincome <- rep(0,50)
for(i in 1:50)
{
  STincome[i] <- mean(subdata1[which(subdata1[,1]==STcode[i]),2])
}

library(plotly)
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
map1 <- cbind(df[,1:2],STincome)
```
```{r}
plot_ly(map1, z = map1[,3], locations = code, type = 'choropleth',
        locationmode = 'USA-states', color = map1[,3], colors = 'Greens',
        marker = list(line = l), colorbar = list(title = "USD"))%>%
  layout(title = '2014 US averge personal income by State', geo = g)
```


#### 1.4.1 RPP map
  Since different states have different cost of living level, we take 2014 RPP (Regional Price Parities) in to consideration. The following maps are associated with RPP and Adjusted Income.
  
```{r,echo=FALSE}
RPP <- c(87.8, 105.7, 96.4, 87.5, 112.4,
         102.0, 108.8, 101.9, 99.1, 92.0,
         116.8, 93.4, 100.7, 91.4, 90.3,
         90.7, 88.7, 91.4, 97.1, 110.3,
         107.1, 94.1, 97.6, 86.7, 89.4,
         94.2, 90.6, 97.7, 105.2, 114.5,
         95.0, 115.7, 91.7, 91.5, 89.3,
         90.1, 99.0, 98.2, 98.7, 90.5,
         88.0, 90.2, 96.6, 97.0, 101.2,
         102.6, 103.8, 88.9, 93.4, 96.2)
map1.7 <- cbind(df[,1:2],RPP)
```
```{r}
plot_ly(map1.7, z = map1.7[,3], locations = code, type = 'choropleth',
        locationmode = 'USA-states', color = map1.7[,3], colors = 'Blues',
        marker = list(line = l), colorbar = list(title = "Index"))%>%
  layout(title = '2014 US RPP by State', geo = g)
```

#### 1.4.2 Adjusted Income Map
```{r,echo=FALSE}
map1.71 <- cbind(df[,1:2],STincome/RPP*100)
```
```{r}
plot_ly(map1.71, z = map1.71[,3], locations = code, type = 'choropleth',
        locationmode = 'USA-states', color = map1.71[,3], colors = 'Oranges',
        marker = list(line = l), colorbar = list(title = "UDS"))%>%
  layout(title = '2014 US adjusted income by State', geo = g)
```


## 2.Employment status in US
Unemployment rate is defined most basically as the percentage of the total labor force that is unemployed but actively seeking employment and willing to work.
That is. Unemployment rate=Unemployed/Total labor force.

### 2.1 Description of Unemployment status
```{r}
us_emp<-nrow(as.matrix(Proj_1[which(ESR==1|ESR==2|ESR==4|ESR==5)]))
us_unemp<-nrow(as.matrix(Proj_1[which(ESR==3)]))
us_rate<- us_unemp/(us_emp+us_unemp)
us_rate
```

### 2.2 Female and Male's Unemployment rate
```{r}
male_emp<-nrow(as.matrix(Proj_1[which(SEX==1&
                                       (ESR==1|ESR==2|ESR==4|ESR==5))]))
male_unemp<-nrow(as.matrix(Proj_1[which(ESR==3&SEX==1)]))
female_emp<-nrow(as.matrix(Proj_1[which(SEX==2&
                                       (ESR==1|ESR==2|ESR==4|ESR==5))]))
female_unemp<-nrow(as.matrix(Proj_1[which(ESR==3&SEX==2)]))
male_rate<- male_unemp/(male_emp+male_unemp)
female_rate<- female_unemp/(female_emp+female_unemp)
male_rate
female_rate
```


### 2.3  Gender, Education and Unemployment Rate
```{r,echo=FALSE}
edu<-matrix()
edu_emp<-vector()
edu_unemp<-vector()

for (i in 1:24){
  edu_emp[i]<-nrow(as.matrix(Proj_1[which(SCHL==i&(ESR==1|ESR==2|ESR==4|ESR==5))]))
  edu_unemp[i]<-nrow(as.matrix(Proj_1[which(ESR==3&SCHL==i)]))
  edu[i]<-edu_unemp[i]/(edu_emp[i]+edu_unemp[i])
}

edu<-cbind(sort(unique(SCHL)),edu)
colnames(edu)<-c("education attainment","unemployment rate")

edu_atm<-c("a.No schooling completed",
           "b.Nursery school/preschool",
           "c.Kindergarten",
           "d.Grade 1",
           "e.Grade 2",
           "f.Grade 3",
           "g.Grade 4",
           "h.Grade 5",
           "i.Grade 6",
           "j.Grade 7",
           "k.Grade 8",
           "l.Grade 9",
           "m.Grade 10",
           "n.Grade 11",
           "o.Grade 12-no diploma",
           "p.Regular high school diploma",
           "q.GED or alternative credential",
           "r.Some college, but less than 1 year",
           "s.1 or more years of college credit, no degree",
           "t.Associate's degree",
           "u.Bachelor's degree",
           "v.Master's degree",
           "w.Professional degree",
           "x.Doctorate degree")
edu<-as.data.frame(edu)
edu[,1]<-as.character(edu[,1])
for (i in 1:24){
  edu[i,1]<-edu_atm[i]
}

edu_female_emp<-vector()
edu_female_unemp<-vector()
edu_male_emp<-vector()
edu_male_unemp<-vector()
edu_female<-matrix()
edu_male<-matrix()

for (i in 1:24){
  edu_female_emp[i]<-nrow(as.matrix(Proj_1[which(SCHL==i&SEX==2&(ESR==1|ESR==2|ESR==4|ESR==5))]))
  edu_female_unemp[i]<-nrow(as.matrix(Proj_1[which(ESR==3&SCHL==i&SEX==2)]))
  edu_female[i]<-edu_female_unemp[i]/(edu_female_emp[i]+edu_female_unemp[i])
  
  edu_male_emp[i]<-nrow(as.matrix(Proj_1[which(SCHL==i&SEX==1&(ESR==1|ESR==2|ESR==4|ESR==5))]))
  edu_male_unemp[i]<-nrow(as.matrix(Proj_1[which(ESR==3&SCHL==i&SEX==1)]))
  edu_male[i]<-edu_male_unemp[i]/(edu_male_emp[i]+edu_male_unemp[i])
}

edu_female<-as.data.frame(cbind(sort(unique(SCHL)),edu_female,rep("female",24)))
colnames(edu_female)<-c("education attainment","unemployment rate","filter")

edu_female[,1]<-as.character(edu_female[,1])
for (i in 1:24){
  edu_female[i,1]<-edu_atm[i]
}
edu_female[,2]<-as.numeric(paste(edu_female[,2]))

edu_male<-as.data.frame(cbind(sort(unique(SCHL)),edu_male,rep("male",24)))
colnames(edu_male)<-c("education attainment","unemployment rate","filter")

edu_male[,1]<-as.character(edu_male[,1])
for (i in 1:24){
  edu_male[i,1]<-edu_atm[i]
}
edu_male[,2]<-as.numeric(paste(edu_male[,2]))

edu_all<-as.data.frame(cbind(edu,rep("all",24)))
colnames(edu_all)<-c("education attainment","unemployment rate","filter")

edu_gender<-rbind(edu_female,edu_male,edu_all)

education.attainment<-edu_gender[,1]
unemployment.rate<-edu_gender[,2]
filter<-edu_gender[,3]

library(rbokeh)
```
```{r}
figure(width = NULL, height = NULL) %>%
ly_points( unemployment.rate,education.attainment,data = edu_gender,
    color = filter, glyph = filter,
    hover = list( unemployment.rate,education.attainment))

```
```{r,echo=FALSE}
library(tidyr)
library(plotly)
s_Women<-as.numeric(paste(edu_female[,2]))
s_Men<-as.numeric(paste(edu_male[,2]))
edu_gender_sub<-as.data.frame(cbind(edu_atm,s_Women,s_Men))
```
```{r}
gather(edu_gender_sub, Sex, value, s_Women, s_Men) %>%
  plot_ly(x = value, y = edu_atm, mode = "markers",
          color = Sex, colors = c("pink", "blue")) %>%
  add_trace(x = value, y = edu_atm, mode = "lines",
            group = edu_atm, showlegend = F, line = list(color = "gray")) %>%
  layout(
    title = "Gender earnings disparity",
    xaxis = list(title = "Annual Salary (in thousands)"),
    margin = list(l = 65)
  )
```


###2.4 States and Unemployment Rate
```{r,echo=FALSE}
STcode <- c(1,2,4,5,6,8,9,10,12,13,15:42,44:51,53:56)
ST_emp <-vector()
ST_unemp<-vector()
ST_unemp_rate<-vector()

for(i in 1:50)
{
  ST_emp[i] <- nrow(as.matrix(Proj_1[which(ST==STcode[i]&(ESR==1|ESR==2|ESR==4|ESR==5))]))
  ST_unemp[i] <- nrow(as.matrix(Proj_1[which(ST==STcode[i]&ESR==3)]))
  ST_unemp_rate[i]<-ST_unemp[i] /(ST_emp[i]+ST_unemp[i])
}

###plot
library(plotly)
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
map1 <- cbind(df[,1:2],ST_unemp_rate)
```
```{r}
plot_ly(map1, z = map1[,3], locations = code, type = 'choropleth',
        locationmode = 'USA-states', color = map1[,3], colors = 'Greens',
        marker = list(line = l), colorbar = list(title = "Unemp_ratae"))%>%
  layout(title = '2014 US unemployment rate by State', geo = g)
```


### 2.5 Education degree and Employment status
```{r,echo=FALSE}
library(networkD3)
netfile<-matrix(0,nrow=144,ncol=3)
colnames(netfile)<-c("source","target","value")
netfile[,1]<-sort(rep(0:23,6))
netfile[,2]<-rep(24:29)

for (i in 1:144){
  netfile[i,3]<-nrow(as.matrix(Proj_1[which(SCHL==(netfile[i,1]+1)&
                                            ESR==(netfile[i,2]-23))]))
}

nodes<-as.data.frame(matrix(0,nrow=30,ncol=2))
colnames(nodes)<-c("name","group")
nodes[,1]<-c("a.No schooling completed",
           "b.Nursery school/preschool",
           "c.Kindergarten",
           "d.Grade 1",
           "e.Grade 2",
           "f.Grade 3",
           "g.Grade 4",
           "h.Grade 5",
           "i.Grade 6",
           "j.Grade 7",
           "k.Grade 8",
           "l.Grade 9",
           "m.Grade 10",
           "n.Grade 11",
           "o.Grade 12-no diploma",
           "p.Regular high school diploma",
           "q.GED or alternative credential",
           "r.Some college, but less than 1 year",
           "s.1 or more years of college credit, no degree",
           "t.Associate's degree",
           "u.Bachelor's degree",
           "v.Master's degree",
           "w.Professional degree",
           "x.Doctorate degree",
           "Civilian employed, at work",
           "Civilian employed, with a job",
           "Unemployed",
           "Armed forces, at work",
           "Armed forces, with a job",
           "Not in labor force")

nodes[,2]<-rep(1:30)

netfile<-as.data.frame(netfile)
```
```{r}
sankeyNetwork(Links = netfile, Nodes = nodes, Source = "source",
Target = "target", Value = "value", NodeID = "name",
units = "TWh", fontSize = 12, nodeWidth = 30)
```
```{r,echo=FALSE}
netfile1<-netfile
netfile1[,3]<-netfile[,3]/10000
```
```{r}
forceNetwork(Links = netfile1, Nodes = nodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             radiusCalculation = "Math.sqrt(d.nodesize)+6", 
             Group = "group", opacity = 0.6,opacityNoHover = TRUE
             )
```


