---
title: "ADS_Project1"
author: "ying"
date: "9/16/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###We first select the variables which we would use in the analysis.
####Include:

   Variables     | Dictionary
   ------------- | -------------
1. SERIALNO      | Unique code for person
2. ST            | State Code
3. SEX           | Gender
4. AGEP          | Age
5. SCHL          | Educational Attainment
6. FOD1P         | Field of Degree - First Entry
7. FOD1P         | Field of Degree - Second Entry
8. WKHP          | Hours Worked per Week
9. WAGP          | Wages or salary income past 12 months
10.ESR           | Employment Status Recode
11.PINCP         | Total Person's Income
12.PERCP         | Total Person's Earning
13.JWMNP         | Travel time to work

###Here we merge the a part and b part data into one dataset:Proj_1

```{r}
setwd("~/Desktop/csv_pus")
Origdata1<-read.csv("ss14pusa.csv",header=T)
Proj_1a<-cbind(Origdata1$SERIALNO,Origdata1$ST,Origdata1$SEX,
               Origdata1$AGEP,Origdata1$SCHL,Origdata1$FOD1P,Origdata1$FOD2P,
               Origdata1$WKHP,Origdata1$WAGP,Origdata1$ESR,Origdata1$PINCP,
               Origdata1$PERNP,Origdata1$JWMNP)
Origdata2<-read.csv("ss14pusb.csv",header=T)
Proj_1b<-cbind(Origdata2$SERIALNO,Origdata2$ST,Origdata2$SEX,
               Origdata2$AGEP,Origdata2$SCHL,Origdata2$FOD1P,Origdata2$FOD2P,
               Origdata2$WKHP,Origdata1$WAGP,Origdata2$ESR,Origdata2$PINCP,
               Origdata2$PERNP,Origdata1$JWMNP)
Proj_1<-rbind(Proj_1a,Proj_1b)
```



Here we have to read the dictionary carefully since there are NA in the columns are actually have means.
The meanings in the NA in different variables:

Variables     | NA meanings
------------- | -------------
SERIALNO      | No NA
ST            | No NA
SEX           | No NA
AGEP          | No NA
SCHL          | Less than 3 years old
FOD1P         | Less than bachelor's degree
FOD1P         | Less than bachelor's degree
WKHP          | Less than 16 years old/did not work
WAGP          | Less than 15 years old
ESR           | Less than 16 years old
PINCP         | Less than 15 years old
PERNP         | Less than 15 years old
JWMNP         | not a worker or worker who worked at home

Therefore in the next analysis, we have to pay attention to  the data meanings.

1.What influences people's income
  +1.1 Description of income
  +1.2 Female and Male's Income.
  +1.3 Age and Income
  +1.4 Education and Income
  +1.5 Occupation and Income
  +1.6 Work Hour and Income
  +1.7 Travel Work Time and Income
  +1.8 States and Income

2.Employment status in the US
  +2.1 Description of Employment status
  +2.2 Female and Male's Employment status
  +2.3 Education and Employment status
  +2.4 Occupation and Employment status
  +2.5 States and Employment status

3.Are incomes and employment rate in the direct proportion?
  +3.1 Age-Income-Employment rate
  +3.2 Gender-Income-Employment rate
  +3.3 Education-Income-Employment rate
  +3.4 Occupation-Income-Employment rate
  +3.5 States-Income-Employment rate

## 1.8 States map of average income
```{r}
data <- Proj_1
subdata1 <- data[,c(2,11)]
subdata1 <- subdata1[which(subdata1[,2]>0),]
STcode <- c(1,2,4,5,6,8,9,10,12,13,15:42,44:51,53:56)

STincome <- rep(0,50)
for(i in 1:50)
{
  STincome[i] <- mean(subdata1[which(subdata1[,1]==STcode[i]),2])
}

library(plotly)
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
map1 <- cbind(df[,1:2],STincome)

plot_ly(map1, z = map1[,3], locations = code, type = 'choropleth',
        locationmode = 'USA-states', color = map1[,3], colors = 'Greens',
        marker = list(line = l), colorbar = list(title = "USD"))%>%
  layout(title = '2014 US averge personal income by State', geo = g)
```

## States map of weekly working hours

```{r}
subdata3 <- data[,c(2,8)]
subdata3 <- subdata3[which(subdata3[,2]>0),]
STworkt <- rep(0,50)
for(i in 1:50)
{
  STworkt[i] <- mean(subdata3[which(subdata3[,1]==STcode[i]),2])
}

library(plotly)
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
map2 <- cbind(df[,1:2],STworkt)

plot_ly(map2, z = map2[,3], locations = code, type = 'choropleth',
        locationmode = 'USA-states', color = map2[,3], colors = 'Reds',
        marker = list(line = l), colorbar = list(title = "HOURS"))%>%
  layout(title = '2014 US averge personal working time by State', geo = g)
```


## 3D plot on income vs education level and working time
## without gender
```{r}
subdata2<- data[,c(5,8,11)]
subdata2 <- subdata2[which(rowSums(subdata2)>0),]
max(subdata2[,2])
max(subdata2[,1])
min(subdata2[,2])
min(subdata2[,1])
income_mat <- matrix(0,nrow=8,ncol=10)
for(i in 1:8){
  for(j in 1:10){
    index <- which(subdata2[,1]>3*i-3 & subdata2[,1]<3*i+1 & subdata2[,2]>10*j-10 & subdata2[,2]<10*j+1)
    income_mat[i,j] <- mean(subdata2[index,3])
  }
}
plot_ly(z=income_mat,type="surface")
```

## with gender (2 level plot)
```{r}
subdata4<- data[,c(5,8,11,3)]
subdata4 <- subdata4[which(rowSums(subdata4)>0),]
max(subdata4[,2])
max(subdata4[,1])
min(subdata4[,2])
min(subdata4[,1])
income_mat_male <- matrix(0,nrow=8,ncol=10)
income_mat_female <- matrix(0,nrow=8,ncol=10)
for(i in 1:8){
  for(j in 1:10){
    index <- which(subdata4[,1]>3*i-3 & subdata4[,1]<3*i+1 &
                    subdata4[,2]>10*j-10 & subdata4[,2]<10*j+1 &
                     subdata4[,4]==1)
    income_mat_male[i,j] <- mean(subdata4[index,3])
  }
}
for(i in 1:8){
  for(j in 1:10){
    index <- which(subdata4[,1]>3*i-3 & subdata4[,1]<3*i+1 &
                     subdata4[,2]>10*j-10 & subdata4[,2]<10*j+1 &
                     subdata4[,4]==2)
    income_mat_female[i,j] <- mean(subdata4[index,3])
  }
}

plot_ly(z=income_mat_male,type="surface",showscale=FALSE) %>%
  add_trace(z=income_mat_female, type="surface", showscale=FALSE, opacity=0.98)
```

## 3D plot on income vs age and working time
```{r}
subdata2<- data[,c(4,8,11)]
subdata2 <- subdata2[which(rowSums(subdata2)>0),]
max(subdata2[,2])
max(subdata2[,1])
min(subdata2[,2])
min(subdata2[,1])
income_mat <- matrix(0,nrow=17,ncol=10)
for(i in 1:17){
  for(j in 1:10){
    index <- which(subdata2[,1]>5*i+10 & subdata2[,1]<5*i+16 & subdata2[,2]>10*j-10 & subdata2[,2]<10*j+1)
    income_mat[i,j] <- mean(subdata2[index,3])
  }
}
plot_ly(z=income_mat,type="surface")
```


